---
title: "Initial Korean covid 19"
author: "Filipe"
date: "October 10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Methods
result ~ categorical 
sex ~ categorical 

since result and sex are both categorical we should use the cat~cat helper file. 

$H_0$ Dying is independent of sex.    
$H_A$ Dying is dependent of sex.     



## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

the barchart shows that males die at a higher percentage than females. 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

among the people who died 64% were males and 36% were females. 2:1

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

p value for the chi-squared test = 0.001 so we reject the null hypothesis so that means there is a difference between the two just by looking at the numbers. the fisher's exact test odd ratio 2.3 so males are 2.3X more likely to die to covid than women. 

## Conclusions

Is dying independent of sex? Because of the barcharts and tables, males are more likley to die than females according to the data. the data was from south korea 2022. so we can conclude that deaths from covid are dependent on sex. p value for the chi-squared test = 0.001